A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-05 paper.tex
reply to some comments

The semantics of these cast representations are defined in many
different papers with separate reduction rules, which makes it
difficult to isolate the key differences between them.  In this paper
we present the first set of reduction rules for a cast calculus that
is parameterized over the cast representation
(Section~\ref{sec:framework}).
\todo[inline]{JS: First? What about \url{https://arxiv.org/abs/2001.11560}?}
The semantics takes the form of a CEK
abstract machine and it formulates the interface between the machine
and the different cast representations as an abstract data type,
namely the \emph{Cast ADT}.

Of course, a space-efficient cast representation must be proved
correct. That is, a language implementation that uses space-efficient
casts should give the same output as the language semantics. Our
generic CEK machine comes with generic theorems
(Section~\ref{sec:framework}) that researchers can use to prove a cast
representation correct.  We believe this framework can reduce the
burden of proving correctness theorems and help people focus on the
cast representations.
\todo[inline]{JS: Say the framework is in Agda.}
For illustration, we apply the framework to
Lazy UD coercions in normal form~\citep{Siek:2015:BCT:2737924.2737968}
and Lazy UD threesomes~\citep{Siek:2010:TWB:1706299.1706342}. We
believe the framework is also applicable to other cast
representations.  The framework and illustrations are available at the
following URL:
\[\text{\urlFrameworkAndIllustration}\]

@@

Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\RMachine{B}) and Fig.~\ref{fig:R-machine-syntax} gives a grammar for machine 
states $s$, including a definition of values $v$ and value typing.
\todo[inline]{JS: explain the name choice of $\mathcal{R}$.}
%
This machine is parameterized over blame strategy $B$, which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions: those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.
That is, the machine state is either in an evaluating
$\sexpr{e}{E}{K}$ or returning $\scont{v}{K}$ configuration, 
respectively. Additionally, there are two halting configurations: \shalt{v} 
halts with a value, while \error{l} halts with a blame label. 
Observables include all value constructors and blame labels, as in 
\citep{siek2012interpretations}. 
The function converting values to observables ($observe(v) = o$) is defined in 
the obvious way.

>>>>

The semantics of these cast representations are defined in many
different papers with separate reduction rules, which makes it
difficult to isolate the key differences between them.  In this paper
we present a set of reduction rules for a cast calculus that
is parameterized over the cast representation
(Section~\ref{sec:framework}).
\todo[inline]{JS: First? What about \url{https://arxiv.org/abs/2001.11560}?}
\todo[inline]{KC: My bad. I don't remember how this word gets in but I
definitely should be more careful about these kind of words. BTW, should
the arxiv paper go into Related Work?}
The semantics takes the form of a CEK
abstract machine and it formulates the interface between the machine
and the different cast representations as an abstract data type,
namely the \emph{Cast ADT}.

Of course, a space-efficient cast representation must be proved
correct. That is, a language implementation that uses space-efficient
casts should give the same output as the language semantics. Our
generic CEK machine comes with generic theorems
(Section~\ref{sec:framework}) that researchers can use to prove a cast
representation correct.  We believe this framework can reduce the
burden of proving correctness theorems and help people focus on the
cast representations.
\todo[inline]{JS: Say the framework is in Agda.}
\todo[inline]{KC: Has said, at the end of this paragraph}
For illustration, we apply the framework to
Lazy UD coercions in normal form~\citep{Siek:2015:BCT:2737924.2737968}
and Lazy UD threesomes~\citep{Siek:2010:TWB:1706299.1706342}. We
believe the framework is also applicable to other cast
representations.  The framework and illustrations (written in Agda)
are available at the following URL:
\[\text{\urlFrameworkAndIllustration}\]

@@

Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\RMachine{B}) and Fig.~\ref{fig:R-machine-syntax} gives a grammar for machine 
states $s$, including a definition of values $v$ and value typing.
\todo[inline]{JS: explain the name choice of $\mathcal{R}$.}
\todo[inline]{KC: I want to space-efficient machine to be S, so ... B might be a better name 
because it reminds people of the $\lambda B$ calculus.
Or $T$ for type-based cast?
Or $C$ for cast-calculus?
Or $D$ for definitional semantic?
How do you think?}
%
This machine is parameterized over blame strategy $B$, which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
\todo[inline]{KC: Technically a $B$ should define a predicate over pretypes, which specifies
what is injectable, and a apply-cast function. Of course, the set of values that the apply-cast
function works with depends on the set of injectable types.
\url{https://github.com/Gradual-Typing/Equivalence-of-Cast-Calculi/blob/master/equivalence-of-cast-calculi/R/BlameStrategies.agda}
Do you want me to put these contraints in the paper?}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions: those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.
That is, the machine state is either in an evaluating
$\sexpr{e}{E}{K}$ or returning $\scont{v}{K}$ configuration, 
respectively. Additionally, there are two halting configurations: \shalt{v} 
halts with a value, while \error{l} halts with a blame label. 
Observables include all value constructors and blame labels, as in 
\citep{siek2012interpretations}. 
The function converting values to observables ($observe(v) = o$) is defined in 
the obvious way.
